---
title: "ggplot2 Worksheet"
output: html_document
authors: Azka Javaid & Pei Gong
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
require(mosaic)
require(ggplot2)
#install.packages("RColorBrewer")
require(RColorBrewer)
```

##Aesthetics (size, shape, alpha)

```{r}
#Point shape, size and alpha (controls point shading intensity) can be specified in geom_point
#Use shapes 21-25 to see the fill attribute 
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, fill = Species)) + geom_point(shape = 21, size = 3, alpha = 0.3)
```

#### Try different specifications of the alpha level as well as shape and size. Complete list of possible shapes can be obtained from online R documentation and is also available in the PowerPoint presented. 

##Aesthetics (jitter specification)

```{r}
posn.j <- position_jitter(width = 0.5) #width can be changed depending on the data scatter
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) + geom_point(position = posn.j)
```

#### Change values of jitter command to visualize the scatter dispersal in the iris dataset. 

##Aesthetics (axis labels and legends)

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) + geom_point(position = "jitter") +
  scale_x_continuous("Length", limits = c(2,8), breaks = seq(2,8,3)) + 
  scale_y_continuous("Width", limits = c(2,4), breaks = seq(2,4,0.5)) +
  scale_color_discrete("Species", labels = c("Setosa", "Versicolour","Virginica"))
```

##Aesthetics (color and RColorBrewer)

```{r}
#display.brewer.all() #displays all possible brewer color combinations 
ggplot(iris, aes(x = Sepal.Length)) +
geom_bar(aes(fill = Species))+scale_fill_brewer(palette = "Reds")
```

#### Uncomment display.brewer.all() and replace palette with another color combination to see how the plot changes. 

##Geometries (position and bar plots)

```{r}
barplot1 <- ggplot(mtcars, aes(x = factor(cyl), fill = factor(am)))
#Default specified stack (uses counts)
barplot1 + geom_bar(position = "stack")

#Fill accounts for proportions
barplot1 + geom_bar(position = "fill")  

#Dodging allows one to separate each factor's affect in a distinguished bar (uses counts)
barplot1 + geom_bar(position = "dodge")
```

##Geometries (boxplot)

```{r}
#Specified binwidth can be altered (smaller binwidth: more detailed representation of the data)
ggplot(iris, aes(x = Species, y = Sepal.Width)) + geom_boxplot()
```

#### Change the y attribute to Sepal.Width to visualize the width for each specie and compare with the plot above.

##Facets

````{r}
base <- ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point()

#Separate rows by transmission type (am)
base + facet_grid(am ~ .)

#Separate columns by transmission type (am)
base + facet_grid(. ~ am)
```

####Explore the two different methods of faceting the data presented above with the cyl attribute in the mtcars dataset. 

##Statistics (Group and Individual Effects)

```{r}
#Individual effect of each level of the cyl attribute
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point() + stat_smooth(method = "lm", se = F)

#Group effect of the cyl attribute
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl), group = 1)) + geom_point() + stat_smooth(method = "lm", se = F)

#Overlaying the individual and the group effects
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point() + stat_smooth(method = "lm", se = F) +
  stat_smooth(method = "lm", se = F, aes(group = 1))
```

##Statistics (Confidence Intervals)

```{r}
#Incorporation of the cyl attribute with the confidence interval cloud
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point() + stat_smooth(method = "lm")
```


##Theme (Changing location of the legend)
```{r}
base <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) + geom_point(position = posn.j)
base + theme(legend.position = c(0.89, 0.86)) #manually specify the location 
base + theme(legend.position = "bottom") 
base + theme(legend.direction = "horizontal")
```

##Theme (Alternate representations of the xyplot by Species)

```{r}
base <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) + geom_point(position = posn.j)
#Changing the background color and other attributes of the base plot (same plot as in the presentation)

theme1 <- theme(panel.background = element_blank(), 
                   legend.background = element_blank(),
                   legend.key = element_blank(),
                   panel.grid = element_blank(),
                   axis.text = element_text(colour = "black"),
                   axis.line = element_line(colour = "black")
                   )
base + theme1
```


####Change the colour in the element_text to observe the effect on the graph.

####Experiment with the theme above, create another theme of choice and apply it to the base plot. 
